Length Gauge Theory of Surface Second Harmonic Generation

### 2014-11-16 paper-v-0.tex
bms-1

The unperturbed Hamiltonian 
is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
Functional Theory (DFT), for convenience usually within the Local  
Density Approximation (LDA). Thus, we labeled the hamiltonian with the  
superscript LDA for convenience, although we could use any other
approximation, like GGA, and our derivation would remain the same. Then,
\begin{equation}\label{h0}
H^{\mathrm{LDA}}_{0}=\frac{p^{2}}{2m_e}+V
\end{equation}
with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
$V$
 the periodic crystal potential, where we neglect spin-orbit terms.
To be more general in our derivation of
$\chi^{abc}(-2\omega;\omega,\omega)$, we asume
that the crystal potential could have a spatially non-local
contribution, as is customary for most
pseudopotentials, and then we replace $V$ with
\begin{align}\label{ps}
V^{\mathrm{ps}}=V(\bfr)+V^{\mathrm{nl}}
,
\end{align}
where 
$V(\bfr)$ and $V^{\mathrm{nl}}$  are the local and 
non-local contributions to the pseudopotential $V^{\mathrm{ps}}$,
 respectively.
Now, the 
Schr\"odinger equation reads 
\begin{align}\label{ache.4} 
\left(
\frac{-\hbar^2}{2m_e}\nabla^2 
 +  V(\bfr)\right)\psi_{n\bfk}(\bfr) 
 + \int d\bfr' V^\nl(\bfr,\bfr')\psi_{n\bfk}(\bfr') 
=\hbar \omega_i\psi_{n\bfk}(\bfr) 
,
\end{align} 
where
$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$,
and
$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
is cell periodic. 
In case of
a local potential, i.e. $V= V(\bfr)$, 
like that of all-electron schemes, 
we simply omit the contribution coming
from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
the results that we will derive.

As is well known, the use of the solutions of Eq.~\eqref{ache.4} as single 
particle states leads to an underestimation of the band gap. A standard
procedure to correct for this is to
use the so-called ``scissors approximation'', by which one rigidly shifts the
conduction bands in energy so that the band gap corresponds to the accepted
experimental band gap; this is often in fairly good agreement with the GW
band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
Concurrently, one uses the LDA wave functions, since they produce band
structures with dispersion relations similar to those predicted by the GW
approximation. Mathematically, one adds the scissors (non-local) term 
$S$, to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$, i.e. 
\begin{equation*}
H^\sigma_{0}=H^{\mathrm{LDA}}_{0}+S
\end{equation*}
where 
\begin{equation}
S=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
|n\mathbf{k}\rangle\langle n\mathbf{k}|,
\label{hats}
\end{equation}
with $\hbar \sigma$  the rigid ($\mathbf{k}$-independent) energy correction to be
applied. 
%Several properties of $S(\mathbf{r},\mathbf{p})$ are shown in the
%appendix.
The unscissored and scissored Hamiltonians satisfy 
\begin{eqnarray*}
H^{\mathrm{LDA}}_{0}\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\label{hamils} \\
H_{0}^\sigma\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\end{eqnarray*}
where the scissor-shifted energies, 
$\omega_{n}^\sigma(\mathbf{k})$ , are given by
\begin{equation}
\omega_{n}^\sigma(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\sigma.
\label{wese}
\end{equation}
We emphasize that the scissored
Hamiltonian has the same eigenfunctions as the unscissored Hamiltonian.

@@

Therefore,
taking the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
Eq.~\eqref{rde}, we obtain 
$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\sigma_{nm}-\tilde\omega)t}$,
with 
\begin{equation}\label{j.1}
B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\sigma_{nm}-\tilde\omega},
\end{equation}
and 
\begin{eqnarray}\label{j.2}
(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\sigma_{nm}-2\tilde\omega}\bigg[
i\sum_{\ell }\Big(r_{n\ell }^{b}B_{\ell m}^{c}-B_{n\ell }^{c}r_{\ell m}^{b}
\Big)  
-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\sigma_{nm}-2\tilde\omega)t}
.
\end{eqnarray}
We have used the fact that for a cold semiconductor $\partial
f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
term vanishes identically. 
%In Appendix \ref{gender} we show how the generalized derivative of 
%$(B^a_{nm})_{;k^b}$ can be obtained.
Notice that in Eq.~\eqref{j.2} all the indices are different from each
other since due to the $f_{nm}$ factor in Eq.~\eqref{j.1},
$B^a_{nn}=0$. The dependence of $\bfk$ of all quantities is implicitly understood from
here on. 
Above expression will be used in the next subsection.

@@

The approach we use to study the surface of a semi-infinite
semiconductor crystal is as follows. Instead of using a
semi-infinite system, we replace it by a super slab that consist of a finite
slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
super slab, which constitutes the unit cell,
is repeated to form a full three dimensional crystalline structure.
The slab itself consists of a front and back surface and sub-surface regions, and in between these
a region that is equivalent to
bulk of the system. 
In general the surface of a crystal reconstructs or relaxes as the atoms
move to find equilibrium positions. This is due to the fact that
the otherwise
balanced forces are disrupted when the surface atoms do not find their 
partner atoms that are now absent at the surface of the slab. 
To take the reconstruction or relaxation into account, we take ``surface'' to mean
the true surface of the first layer of atoms, and
some of the atomic sub-layers adjacent to it.
Since the front and the back
surfaces of the slab are usually identical the total slab is
centrosymmetric. This would imply that 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
find a scheme 
in order to have a finite $\chi_{\rma\rmb\rmc}$ representative of the
surface. Even if the front and back surfaces of the slab 
are different, breaking the centrosymmetry and therefore giving an
overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
and the back surface $\chi^b_{\rma\rmb\rmc}$ from the slab
susceptibility 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$.
We have omitted the frequency arguments of $\chi_{abc}$ for notation convenience.

A convenient way to accomplish the separation of the SH signal of
either surface is to introduce a ``cut function'', $\calc(z)$, which is 
usually taken to be unity over one half of the slab and zero over 
the other half.\cite{reining_microscopic_1994}
 In this case $\calc(z)$ will give the contribution of the 
side of the slab for which $\calc(z)=1$. 
As was done for the linear response,\cite{mendozaPRB06}
we can generalize this 
simple choice for $\calc(z)$ by a top-hat cut function
$\calc^\ell(z)$ that selects a given layer,
\begin{equation}
\label{sz}
\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{\rmb})  
            \Theta(z_\ell-z+\Delta_\ell^f),
\end{equation} 
where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
is the distance that the $\ell$-th layer extends towards the front
$(f)$ or back $(b)$ from its $z_\ell$ position. 
 We take $z_\ell$
 at the center of an atom that 
belongs to layer $\ell$, and thus above equation would give the $\ell$-th 
 atomic-layer 
contribution to the non-linear optical response.
$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
(see Fig.~\ref{fslab}).
\begin{figure}[b]
\centering
%\includegraphics[height=5cm,width=7cm]{slab}
\includegraphics[scale=.7]{images/slab}
\caption{A sketch of a slab where the circles represent atoms.\label{fslab}}
\end{figure}

@@

 \equiv \mathbf{J}^{(N),\ell}(t),

@@

where $\mathbf{J}^{(N),\ell}(t)$ is the current of the

@@

e{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}(\mathbf{k})

@@

\bfJ^{(N),\ell}(2\go)=\frac{e}{\gO}

@@

\calbv^{\gs,\ell}_{mn}(\mathbf{k}) 

@@

{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}({\mathbf k})

@@

\calbv^{\gs,\ell}

@@

\calbv^{\lda,\ell}

@@

\calbv^{S,\ell}

@@

\calbv^{\nl,\ell}
\frac{1}{m_e}
\calbp^{\ell}
\calbv^{\nl,\ell}

@@

$\calbv^{\sigma,\ell}_{nm}(\bfk)$   

@@

$\calbv^{\lda,\ell}_{nm}(\bfk)$,

@@

$\calbv^{\sigma,\ell}_{nm}(\bfk)\ne

@@

\calbv^{\lda,\ell}_{nm}(\bfk)$,

@@

$\calbv^{\lda,\ell}$ and $\calbv^{S,\ell}$

@@

To limit the SHG response to one surface, the equivalent of Eq.~\eqref{nl.4} 
for $\calbv^\ell=\calbp^\ell/m_e$ was proposed in 
Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
\onlinecite{mendozaPRL98},
\onlinecite{mendoza_ab_2001},
\onlinecite{sanoPRB02},
 and \onlinecite{mejia_layer-by-layer_2004}. 
The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003,castilloPRB03,mottaCMS10} 
used the equivalent Eq.~\eqref{sz}, 
%limiting the current response
%to a particular layer of the slab and used 
to obtain the
anisotropic linear optical response of semiconductor surfaces.
However, the first formal derivation
of this scheme 
for the linear response 
is presented in
Ref.~\onlinecite{mendozaPRB06}, 
and here in this 
article, for the second harmonic optical response of semiconductors.

@@

\bfP^{(2),\ell}(2\go)

@@

\frac{\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})}

@@

+i\sum_\ell\left(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) -
  B_{n\ell}^{\rmc}(\bfk,\go) 
  r_{\ell m}^{\rmb}\right)

@@

$\chi^{\ell}_{\rma\rmb\rmc}(-2\go;\go,\go)$ 

@@

@@

\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}

@@

\sum_{\ell mn}
\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}

@@

\frac{r_{n\ell}^{\rmc} r_{\ell m}^{\rmb} 
f_{m\ell}}{\go^\gs_{\ell m}-\got}
-\frac{r_{n\ell}^{\rmb} r_{\ell m}^{\rmc} 
f_{\ell n}}{\go^\gs_{n \ell}-\got}

@@

The generalized derivative in Eq.~\eqref{chii} is dealt with by using the chain rule 

@@

remove the square in the denominator of the second term in Eq.~\eqref{gene2}, 

@@

\input{appendices/chis}
\noindent where we have splited the interband and intraband $1\go$ and $2\go$

@@

We calculate the nonlinear surface susceptibility as 
\begin{equation}\label{chiijksur}
\bfgchi(-2\go;\go,\go)
= \sum_{\{\ell\}}\bfgchi^{\ell}(-2\go;\go,\go),
where $\{\ell\}$
is meant to be set of layers that one choses. For instance, one go
from 
a layer that encompasses half of the slab  or each atomic layer till
the middle of the slab. In the first case, there is one single summand
in Eq.~\eqref{chiijksur}, in the second case a sum from
$\ell=1$ which denotes
the first layer right at the surface, 
till $\ell=N$ that corresponds to the layer at the middle of the salb, which ought to be
a bulk-like layer.
%that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
%For a centrosymmetric slab, 
%$\bfgchi^{\ell=N}=0$;
We reamark that the value of 
$N$ is not universal, and  
the slab needs to have enough atomic layers for 
%$\bfgchi^{\ell=N}=0$;
%to be satisfied and to 
in order to give converged results for $\bfgchi (-2\go;\go,\go)$. 
We can use Eq.~\eqref{chiijksur} for 
either the front or the back surface. 
\textcolor{red}{Talk about the correct value of $d$}

@@

%%%%%%%%%%%%%
\appendix 
\section{}
We give explicit expressions for the quantities used in the evaluation 
of Eq.~\eqref{calvimchiewn}-\eqref{calvimchi2wn}; when appropriate, some 
intermediate steps are given for their derivation. 
\input{appendices/appendix-vc}%calpcalc G

\input{appendices/appendix-gen_der_r_nonlocal}%gdernl H

\input{appendices/appendix-vnl}%gdernl H

\input{appendices/appendix-tnm}%calt F

\input{appendices/appendix-gen_der_calv}%calvs C

>>>>

The unperturbed Hamiltonian 
is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
Functional Theory (DFT), for convenience usually within the Local  
Density Approximation (LDA). Thus, we labeled the hamiltonian with the  
superscript LDA for convenience, although we could use any other
approximation, like GGA, and our derivation would remain the same. Then,
\begin{equation}\label{h0}
H^{\mathrm{LDA}}_{0}=\frac{p^{2}}{2m_e}+V
\end{equation}
with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
$V$
 the periodic crystal potential, where we neglect spin-orbit terms.
To be more general in our derivation of
$\chi^{abc}(-2\omega;\omega,\omega)$, we asume
that the crystal potential could have a spatially non-local
contribution, as is customary for most
pseudopotentials, and then we replace $V$ with
\begin{align}\label{ps}
V^{\mathrm{ps}}=V(\bfr)+V^{\mathrm{nl}}
,
\end{align}
where 
$V(\bfr)$ and $V^{\mathrm{nl}}$  are the local and 
non-local contributions to the pseudopotential $V^{\mathrm{ps}}$,
 respectively.
In case of
a local potential, i.e. $V= V(\bfr)$, 
like that of all-electron schemes, 
we simply omit the contribution coming
from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
the results that we will derive.

As is well known, the use of the LDA scheme 
leads to an underestimation of the band gap. A standard
procedure to correct for this is to
use the so-called ``scissors approximation'', by which one rigidly shifts the
conduction bands in energy so that the band gap corresponds to the accepted
experimental band gap; this is often in fairly good agreement with the GW
band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
Concurrently, one uses the LDA wave functions, since they produce band
structures with dispersion relations similar to those predicted by the GW
approximation. Mathematically, one adds the scissors (non-local) term 
$S$, to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$, i.e. 
\begin{equation*}
H^\sigma_{0}=H^{\mathrm{LDA}}_{0}+S
\end{equation*}
where 
\begin{equation}
S=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
|n\mathbf{k}\rangle\langle n\mathbf{k}|,
\label{hats}
\end{equation}
with $\hbar \sigma$  the rigid ($\mathbf{k}$-independent) energy correction to be
applied. 
%Several properties of $S(\mathbf{r},\mathbf{p})$ are shown in the
%appendix.
The unscissored and scissored Hamiltonians satisfy 
\begin{eqnarray*}
H^{\mathrm{LDA}}_{0}\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\label{hamils} \\
H_{0}^\sigma\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\end{eqnarray*}
where the scissor-shifted energies, 
$\omega_{n}^\sigma(\mathbf{k})$ , are given by
\begin{equation}
\omega_{n}^\sigma(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\sigma.
\label{wese}
\end{equation}
We emphasize that the scissored
Hamiltonian has the same eigenfunctions as the unscissored
Hamiltonian,
where
$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
is cell periodic.

@@

Therefore,
taking the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
Eq.~\eqref{rde}, we obtain 
$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\sigma_{nm}-\tilde\omega)t}$,
with 
\begin{equation}\label{j.1}
B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\sigma_{nm}-\tilde\omega},
\end{equation}
and 
\begin{eqnarray}\label{j.2}
(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\sigma_{nm}-2\tilde\omega}\bigg[
i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
\Big)  
-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\sigma_{nm}-2\tilde\omega)t}
.
\end{eqnarray}
We have used the fact that for a cold semiconductor $\partial
f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
term vanishes identically. 
%In Appendix \ref{gender} we show how the generalized derivative of 
%$(B^a_{nm})_{;k^b}$ can be obtained.
Notice that in Eq.~\eqref{j.2} all the indices are different from each
other since due to the $f_{nm}$ factor in Eq.~\eqref{j.1},
$B^a_{nn}=0$. The dependence of $\bfk$ of all quantities is implicitly understood from
here on. 
Above expression will be used in the next subsection.

@@

The approach we use to study the surface of a semi-infinite
semiconductor crystal is as follows. Instead of using a
semi-infinite system, we replace it by a super cell that consist of a finite
slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
super cell, which constitutes the unit cell,
is repeated to form a full three dimensional crystalline structure.
The slab itself consists of a front and back surface and sub-surface regions, and in between these
a region that is equivalent to
bulk of the system. 
In general the surface of a crystal reconstructs or relaxes as the atoms
move to find equilibrium positions. This is due to the fact that
the otherwise
balanced forces are disrupted when the surface atoms do not find their 
partner atoms that are now absent at the surface of the slab. 
To take the reconstruction or relaxation into account, we take ``surface'' to mean
the true surface of the first layer of atoms, and
some of the atomic sub-layers adjacent to it.
Since the front and the back
surfaces of the slab are usually identical the total slab is
centrosymmetric. This would imply that 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
find a scheme 
in order to have a finite $\chi_{\rma\rmb\rmc}$ representative of the
surface. Even if the front and back surfaces of the slab 
are different, breaking the centrosymmetry and therefore giving an
overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
and the back surface $\chi^b_{\rma\rmb\rmc}$ from the slab
susceptibility 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$.
We have omitted the frequency arguments of $\chi_{abc}$ for notation convenience.

A convenient way to accomplish the separation of the SH signal of
either surface is to introduce a ``cut function'', $\calc(z)$, which is 
usually taken to be unity over one half of the slab and zero over 
the other half.\cite{reining_microscopic_1994}
 In this case $\calc(z)$ will give the contribution of the 
side of the slab for which $\calc(z)=1$. 
As was done for the linear response,\cite{mendozaPRB06}
we can generalize this 
simple choice for $\calc(z)$ by a top-hat cut function
$\calc^\ell(z)$ that selects a given layer,
\begin{equation}
\label{sz}
\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
            \Theta(z_\ell-z+\Delta_\ell^f),
\end{equation} 
where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
is the distance that the $\ell$-th layer extends towards the front
$(f)$ or back $(b)$ from its $z_\ell$ position. 
 We take $z_\ell$
 at the center of an atom that 
belongs to layer $\ell$, and thus above equation would give the $\ell$-th 
 atomic-layer 
contribution to the non-linear optical response.
$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
(see Fig.~\ref{fslab}).
\begin{figure}[b]
\centering
%\includegraphics[height=5cm,width=7cm]{slab}
\includegraphics[scale=.7]{images/slab}
\caption{A sketch of the super cell, where the slab corresponds to the
  circles representing the atoms of the system.\label{fslab}} 
\end{figure}

@@

 \equiv \mathbf{J}^{\ell,(N)}(t),

@@

where $\mathbf{J}^{\ell,(N)}(t)$ is the current of the

@@

e{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}(\mathbf{k})

@@

\bfJ^{\ell,(N)}(2\go)=\frac{e}{\gO}

@@

\calbv^{\ell,\gs}_{mn}(\mathbf{k}) 

@@

{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}({\mathbf k})

@@

\calbv^{\ell,\gs}

@@

\calbv^{\ell,\lda}

@@

\calbv^{\ell,S}

@@

\calbv^{\ell,\nl}

@@

$\calbv^{\ell,\sigma}_{nm}(\bfk)$   

@@

$\calbv^{\ell,\lda}_{nm}(\bfk)$,

@@

$\calbv^{\ell,\sigma}_{nm}(\bfk)\ne

@@

\calbv^{\ell,\lda}_{nm}(\bfk)$,

@@

$\calbv^{\ell,\lda}$ and $\calbv^{\ell,S}$

@@

To limit the SHG response to one surface, the equivalent of Eq.~\eqref{nl.4} 
for $\calbp^\ell$ was proposed in 
Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
\onlinecite{mendozaPRL98},
\onlinecite{mendoza_ab_2001},
\onlinecite{sanoPRB02},
 and \onlinecite{mejia_layer-by-layer_2004}. 
The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003,castilloPRB03,mottaCMS10} 
used the equivalent Eq.~\eqref{sz}, 
%limiting the current response
%to a particular layer of the slab and used 
to obtain the
anisotropic linear optical response of semiconductor surfaces.
However, the first formal derivation
of this scheme 
for the linear response 
is presented in
Ref.~\onlinecite{mendozaPRB06}, 
and here in this 
article, for the second harmonic optical response of semiconductors.

@@

\bfP^{\ell,(2)}(2\go)

@@

\frac{\calv^{\ell,\gs,\rma}_{mn}(\mathbf{k})}

@@

+i\sum_q\left(r_{nq}^{\rmb}B_{qm}^{\rmc}(\bfk,\go) -
  B_{nq}^{\rmc}(\bfk,\go) 
  r_{qm}^{\rmb}\right)

@@

@@

%\end{document}  

@@

\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}

@@

\sum_{qmn}
\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}

@@

\frac{r_{nq}^{\rmc} r_{qm}^{\rmb} 
f_{mq}}{\go^\gs_{qm}-\got}
-\frac{r_{nq}^{\rmb} r_{qm}^{\rmc} 
f_{qn}}{\go^\gs_{nq}-\got}

@@

The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 

@@

remove the square in the denominator of the second term of Eq.~\eqref{gene2}, 

@@

\begin{subequations}\label{xis}
\begin{equation}\label{calvimchiewn}
\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,\go}]= 
\frac{\pi |e|^3}{2\hbar^2}
\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gs_{cv}}
\left[
\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
{(2\go^\gs_{cv}-\go^\gs_{cq})} 
-\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
{(2\go^\gs_{cv}-\go^\gs_{qv})}
\right]\gd(\go^\gs_{cv}-\go),
\begin{equation}\label{calvimchiwn}
\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,\go}]= 
\frac{\pi\vert e\vert^3}{2\hbar^2}
\sum_{cv}\frac{1}{(\omega^\gs_{cv})^{2}}
\left[
\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
+\frac{\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}} 
\right]\delta(\omega^\gs_{cv}-\omega),
\begin{equation}\label{calvimchie2wn}
\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,2\go}]= 
-\frac{\pi |e|^3}{2\hbar^2}
\sum_{vc}\frac{4}{\omega^\gs_{cv}}
\left[
\sum_{v'\ne
  v}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
{2\go^\gs_{cv'}-\go^\gs_{cv}}
- \sum_{c'\ne
  c}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
{2\go^\gs_{c'v}-\go^\gs_{cv}}
\right]\gd(\go^\gs_{cv}-2\go),
\begin{equation}\label{calvimchi2wn}
\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,2\go}]= 
 \frac{\pi \vert
   e\vert^{3}}{2\hbar^2}
\sum_{vc}\frac{4}{(\omega^\gs_{cv})^{2}}
\left[\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
\right\}\right] -
\frac{2\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}}\right]\delta(\omega^\gs_{cv}-2\omega)
\end{subequations}
 where we have splited the interband and intraband $1\go$ and $2\go$

@@

The various quantities involved in Eqs.~\eqref{xis} are given in
the Appendix \ref{appe}.

Finally, we calculate the nonlinear surface susceptibility as 
\begin{equation}\label{chiijksur}
\bfgchi(-2\go;\go,\go)
= \sum_{\{\ell\}}\bfgchi^{\ell}(-2\go;\go,\go),
\end{equation} 
where $\{\ell\}$
is meant to be set of layers that one choses. For instance, one go
from 
a layer that encompasses half of the slab  or each atomic layer till
the middle of the slab. In the first case, there is one single summand
in Eq.~\eqref{chiijksur}, in the second case a sum from
$\ell=1$ which denotes
the first layer right at the surface, 
till $\ell=N$ that corresponds to the layer at the middle of the salb, which ought to be
a bulk-like layer.
%that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
%For a centrosymmetric slab, 
%$\bfgchi^{\ell=N}=0$;
We reamark that the value of 
$N$ is not universal, and  
the slab needs to have enough atomic layers for 
%$\bfgchi^{\ell=N}=0$;
%to be satisfied and to 
in order to give converged results for $\bfgchi (-2\go;\go,\go)$. 
We can use Eq.~\eqref{chiijksur} for 
either the front or the back surface. 
%\mathrextcolor{red}{Talk about the correct value of $d$}

@@

%%%%%%%%%%%%%
\appendix 
\section{}\label{appe}
We give explicit expressions for the quantities used in the evaluation 
of Eq.~\eqref{xis}; when appropriate, some 
intermediate steps are given for their derivation. 
%%%%
\subsection{ Expressions for 
\texorpdfstring{${\cal V}^{\ell,a}_{nm}(\bfk)$, 
${\cal C}^{\ell}_{nm}(\bfk)$ 
and
$({\cal C}^{\ell}_{nm}(\bfk))_{;\bfk}$
}{Vnm and Cnm}}\label{calpcalc}

Expanding the wave function in plane waves we obtain
\begin{align}\label{eni.1}
\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
,
\end{align}
where $\{\bfG\}$ are the reciprocal basis vectors satisfying
$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
space, and $A_{n\bfk}(\bfG)$ are the expansion coefficients. Using
$m_e\bfv=-i\hbar\bfgnabla$ into the equivalent of Eq.~\eqref{intj}
 we obtain,\cite{mendozaPRB06}
\begin{align}\label{eni.2}
\calbv^\ell_{nm}(\bfk)=
\frac{\hbar}{2m_e}
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
(2\bfk+\bfG+\bfG')
\gd_{\bfG_\parallel \bfG'_\parallel}  
f^\ell(G_\perp-G'_\perp)
,
\end{align}   
where
\begin{align}\label{vnl.9}
f^\ell(g)=\frac{1}{L}\int_{z_\ell-\gD^b_\ell}^{z_\ell+\gD^f_\ell} e^{igz}dz  
 ,
\end{align}
with $f^{\ell*}(g)=f^\ell(-g)$. 
the reciprocal lattice vectors $\bfG$ are decomposed into components
parallel to the surface $\bfG_\parallel$, and perpendicular to the
surface $G_\perp \hat z$, so
that $\bfG = \bfG_\parallel + G_\perp\hat z$.
The double summation over the $\bfG$ vectors can be efficiently done by  
creating a pointer array to identify all the plane-wave coefficients  
associated with the same $G_\parallel$.  
 Likewise we obtain that
\begin{align}\label{eni.4}
\calc^\ell_{nm}(\bfk)=
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
\gd_{\bfG_\parallel \bfG'_\parallel} 
f^\ell(G_\perp-G'_\perp)
.
\end{align}  
If $\calc^\ell(z)=1$, $f^\ell(g)=\gd_{g0}$, and from Eqs.~\eqref{eni.2} and \eqref{eni.4}
we obtain the full-slab or bulk values
$\bfv_{nm}(\bfk)$ and
$\calc^\ell_{nm}(\bfk)=\gd_{nm}$.

Since for any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, using 
Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn}, we obtain
\begin{align} 
(\calc^\ell_{nm})_{;\bfk}
&=
i
\sum_{q} 
 \left(\bfr_{nq}
\calc^\ell_{qm}
-
\calc^\ell_{nq}
\bfr_{qm}
\right) 
+i\bfr_{nm}(\calc^\ell_{mm}-\calc^\ell_{nn}) 
,
\end{align} 
where we remind the reader that $\bfr_{nm}$ are calculated through
Eq.~\eqref{chon.10} for LDA.
